miniKanren with fair search strategies

### 2019-04-10 Paper/paper.tex
very brief conclusion

%%% NOTE %%%
%
% - deadline?
%
%   miniKanren workshop deadline May 15, 2018
%
%%%%%%%%%%%%

%TODO test cases: quine, '(I love you).

%TODO 1. reformat LaTeX (80 chars per line)
%TODO 2. stat quine
%TODO 3. keep writing

@@

% body
\begin{document}

@@

% TODO rewrite, because Seres did provide a solution for possibly finite stream
% The unfairness of disjunction has been noticed by Seres et 
% al~\citet{seres1999algebra}. Their complete 
% and fair search strategy is also named breadth-first search. Their BFS 
% is fair and is similar to ours. However, their Haskell implementation cannot be 
% translated to Scheme directly. This is partially due to the difference in 
% calling convensions of host languages. Besides, their search space are infinite 
% even when no answers exists. This is not feasible in miniKanren, where 
% users can query all answers.

@@

% KC: I changed the cost definition from ``... needed to *find* the answer''
%                                     to ``... needed to *verify* the answer''

@@

% COmmented out: This idea is borrowed from Silvija Seres's work [*].
% KC: Seres didn't phrase it this way.
%   We might need to compare our definition with Seres's one, instead of 
%   saying the idea is borrowed.

@@

We avoid generating same-cost answers at once by expressing BFS with a queue, 
whose elements thunks that return a new stream. Every \texttt{mk-1} stream has 
zero or one thunk, so we cannot manage it with queue in interesting 
way. Therefore we change the representation of immature parts from thunks to 
lists of thunks. As as consequence, we also change the way to combine mature and 
immature part from \texttt{append} to \texttt{cons}.

After applying this two changes, stream representation becomes more 
complicated. It motivates us to set up an interface between stream 
and the rest of miniKanren. Listed in Fig.~\ref{interface} are all functions 
being aware of the stream representation, but \texttt{take-inf} and its helper 
function, which are explained later. The first three functions are 
constructors: 
\texttt{empty-inf} constructs an empty stream; \texttt{unit-mature-inf} 
constructs a stream with one mature solution; \texttt{unit-immature-inf} 
constructs a stream with one thunk. The \texttt{append-inf} in \texttt{mk-3} is 
relatively straightforwared compared with the \texttt{mk-1} version. 
\texttt{append-map-inf} is more tricky on how to construct the new immature 
part. We can follow the approach in mk-0 and mk-1 -- create a new thunk which 
invoke append-map-inf recursively when forced. But then we need to be careful: 
if we construct the thunk when the old immature part is an empty list, the 
resulting stream might be infinitely unproductive. Beside, all solutions of the 
next lowest cost in \texttt{s-inf} must be computed when the thunk is invoked. 
However sometimes only a portion of these solutions is required to answer a 
query. To avoid the trouble and the too-early computation, we choose to create 
a new thunk for every existing thunk. The next four functions are used only by 
\texttt{ifte} and \texttt{once}. Uninterested readers might skip them. 
\texttt{null-inf?} checks whether a stream is exausted. \texttt{mature-inf?} 
checks whether a stream has some mature solutions. \texttt{car-inf} takes the 
first solution out of a mature stream. \texttt{cdr-inf} drops the first 
solution of a mature stream. Finally, \texttt{force-inf} forces an a immature 
stream to do more computation.

@@

>>>>

%%% NOTE %%%
%
%  SUBMISSION DEADLINE: May 15, 2019
%
%%%%%%%%%%%%

@@

\begin{document}
% TODO how about making all disjunction trees complete by filling missing
% leaves with nevero?

@@

@@

@@

% Commented out: This idea is borrowed from Silvija Seres's work [*].
% KC: Seres didn't phrase it this way.
%   We might need to compare our definition with Seres's one, instead of 
%   saying the idea is borrowed.

@@

We avoid generating same-cost answers at once by expressing BFS with a queue, 
whose elements thunks that return a new stream. Every \texttt{mk-1} stream has 
zero or one thunk, so it is uninteresting to manage them with queue . Therefore 
we change the representation of immature parts from thunks to 
lists of thunks. As as consequence, we also change the way to combine mature and 
immature part from \texttt{append} to \texttt{cons}.

After applying this two changes, stream representation becomes more 
complicated. It motivates us to set up an interface between stream 
and the rest of miniKanren. Listed in Fig.~\ref{interface} are all functions 
being aware of the stream representation, but \texttt{take-inf} and its helper 
function, which are explained later.

The first three functions are constructors: \texttt{empty-inf} constructs an 
empty stream; \texttt{unit-mature-inf} constructs a stream with one mature 
solution; \texttt{unit-immature-inf} constructs a stream with one thunk.

The \texttt{append-inf} in \texttt{mk-3} is relatively straightforwared compared 
with the \texttt{mk-1} version. \texttt{append-map-inf} is more tricky on how to 
construct the new immature part. We can follow the approach in mk-0 and mk-1 -- 
create a new thunk which invoke append-map-inf recursively when forced. But then 
we need to be careful: if we construct the thunk when the old immature part is 
an empty list, the resulting stream might be infinitely unproductive. Beside, 
all solutions of the next lowest cost in \texttt{s-inf} must be computed when 
the thunk is invoked. However sometimes only a portion of these solutions is 
required to answer a query. To avoid the trouble and the too-early computation, 
we choose to create a new thunk for every existing thunk.

The next four functions are used only by \texttt{ifte} and \texttt{once}. 
Uninterested readers might skip them. \texttt{null-inf?} checks whether a stream 
is exausted. \texttt{mature-inf?} checks whether a stream has some mature 
solutions. \texttt{car-inf} takes the first solution out of a mature stream. 
\texttt{cdr-inf} drops the first solution of a mature stream. Finally, 
\texttt{force-inf} forces an a immature stream to do more computation.

\texttt{unit-mature-inf}, \texttt{append-map-inf}, \texttt{empty-inf} and 
\texttt{append-inf} form a \emph{MonadPlus}, where they corresponds to 
\texttt{unit}, \texttt{bind}, \texttt{mzero}, and \texttt{mplus} respectively.

@@

We devise a new search strategy, balanced interleaving DFS. The key idea is to 
make disjunction tree balanced. Changing search strategy from iDFS to biDFS is 
not hard: 2 new functions and 1 modified macro.

We also devise breadth-first search, which is similar to Seres's BFS. These two 
BFSs have similar intuition, but whether they are equivalent is not clear yet. 
Our optimization on BFS with queue, however, is new.
